---
title: "p8105_hw2_hh2767.Rmd"
author: "Haoran Hu"
date: "2018-09-30"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readxl)
```
```{r}
nyc_transit = read_csv(file = ".\\data\\NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>% 
  janitor::clean_names() %>% 
  select(line, station_name, station_latitude, station_longitude, route1:route11, entry, vending, entrance_type, ada) %>% 
mutate(entry = ifelse(entry == "NO", FALSE, TRUE))
```
This dataset contains variables that describe NYC transits. The variables are:lines, station names, station locations, the routes that the station serves, entry type, vendings, ADA compliances and so on. So far, I have selected the variables that we need, cleaned the variable names, and turned entry variable from character variable into logical variable. The dimension of the data set is `r nrow(nyc_transit)`* `r ncol(nyc_transit)`. These data are tidy.

```{r}
distinct_stations = distinct(nyc_transit, station_name, route1, route2, route3, route4, route5, route6, route7, route8, route9, route10, route11, .keep_all = TRUE) 
```
There are `r nrow(distinct_stations)` distinct stations. `r sum(distinct_stations$ada)` of the stations are ada compliant. 

```{r}
etr_novending = filter(nyc_transit, vending == "NO")
no_vend_allow_etr = sum(etr_novending$entry)
num_etr_novending = nrow(etr_novending)
```
The proportion of station entrances / exits without vending that allow entrance is `r round(no_vend_allow_etr/num_etr_novending, 3)`.

In the following part, I will reformat data so that route number and route name are distinct variables, and the data will become clean data.
```{r}
tidy_nyc_transit = gather(nyc_transit, key = route_num, value = route_name, route1:route11) %>% 
  filter(!is.na(route_name)) %>% 
  distinct()

station_serve_A = distinct(tidy_nyc_transit,  station_name, .keep_all = TRUE) %>% 
  filter(route_name == "A")
```
The number of distinct station that serve the A train is `r nrow(station_serve_A)`, and among those stations, `r round(sum(station_serve_A$ada)/nrow(station_serve_A),3)`

#Problem 2
```{r}
wheel_trash = readxl::read_excel(".\\data\\HealthyHarborWaterWheelTotals2017-9-26.xlsx", range = "A2:N258") %>% 
  janitor::clean_names() %>% 
  filter(!is.na(dumpster))
wheel_trash$sports_balls = round(wheel_trash$sports_balls)
wheel_trash$sports_balls = as.integer(wheel_trash$sports_balls)
```

```{r}
pptn_2017 = readxl::read_excel(".\\data\\HealthyHarborWaterWheelTotals2017-9-26.xlsx", sheet = "2017 Precipitation") %>% 
  janitor::clean_names() %>% 
  filter(!is.na(precipitation_in)) %>% 
  filter(!is.na(x_1)) %>% 
  mutate(year = 2017)

pptn_2016 = readxl::read_excel(".\\data\\HealthyHarborWaterWheelTotals2017-9-26.xlsx", sheet = "2016 Precipitation") %>% 
  janitor::clean_names() %>% 
  filter(!is.na(precipitation_in)) %>% 
  filter(!is.na(x_1)) %>% 
  mutate(year = 2016)

pptn = bind_rows(pptn_2016, pptn_2017)

pptn = filter(pptn, pptn$precipitation_in != "Month")
pptn$precipitation_in = as.integer(pptn$precipitation_in)
pptn_precipitation_in = tibble(pptn$precipitation_in)
pptn$precipitation_in = apply(pptn_precipitation_in, 2, function(x) month.name[x])
```

